Flabby and injective objects in toposes

### 2018-10-20 paper-flabby-objects.tex
Finished proof that higher direct images are internal sheaf cohomology

We adopt in this section the axiom of choice, so that enough injectives are
available to ensure that this recipe can be carried out. The resulting sheaf
of~$f_*\O_Y$-modules is well-defined in the following sense: Given a further
injective resolution~$0 \to E \to J^\bullet$, there is up to homotopy precisely
one morphism~$I^\bullet \to J^\bullet$ compatible with the identity on~$E$, and
this morphism induces an isomorphism on cohomology.

Higher direct images are pictured as a ``relative'' version of sheaf
cohomology. Due to the result that injectivity of sheaves of modules can be
characterized in the internal language, we can give the following precise
rendering of this slogan.

The over-locale~$Y$ corresponds to a locale~$I(Y)$ internal to~$\Sh(X)$, in
such a way that the category of internal sheaves over this internal locale
coincides with~$\Sh(Y)$~\cite{XXX}; in particular, a given sheaf~$E$ of~$\O_Y$-modules can
be regarded as a sheaf over~$I(Y)$. Under this equivalence, the morphism~$f : Y
\to X$ corresponds to the unique morphism~$I(Y) \to \mathrm{pt}$ to the
internal one-point locale. Hence it makes sense to construct, from the internal
point of view of~$\Sh(X)$, the sheaf cohomology~$H^n(I(Y), E)$ of~$E$.

\begin{thm}The sheaf on~$X$ obtained by evaluating the expression~``$H^n(I(Y),
E)$'' of the internal language of~$\Sh(X)$ coincides with~$R^n
f_*(E)$.\end{thm}

\begin{proof}We recall that the stack semantics is idempotent in the sense
that~$\Sh(X) \models (\Sh(I(Y)) \models \varphi)$ if and only if~$\Sh(Y)
\models \varphi$. Hence a sheaf of modules on~$Y$ is injective if and only if
it is an injective module from the point of view of~$\Sh(Y)$, if and only if it
is ... XXX Since we have assumed the axiom of choice in the metatheory and
since internal and external injectivity coincide by
Theorem~\ref{thm:injectivity-external-internal}, the statement~``any module has
an injective resolution'' holds internally to~$\Sh(Y)$. Therefore the
statement XXX
\end{proof}

>>>>

Assuming the axiom of choice, there are enough injective sheaves of modules so
that this recipe can be carried out. The resulting sheaf
of~$f_*\O_Y$-modules is well-defined in the following sense: Given a further
injective resolution~$0 \to E \to J^\bullet$, there is up to homotopy precisely
one morphism~$I^\bullet \to J^\bullet$ compatible with the identity on~$E$, and
this morphism induces an isomorphism on cohomology.

Higher direct images are pictured as a ``relative'' version of sheaf
cohomology. Due to the result that injectivity of sheaves of modules can be
characterized in the internal language, we can give a precise
rendering of this slogan: We can understand higher direct images as
internal sheaf cohomology.

The details are as follows. The over-locale~$Y$ corresponds to a locale~$I(Y)$ internal to~$\Sh(X)$, in
such a way that the category of internal sheaves over this internal locale
coincides with~$\Sh(Y)$; in particular, a given sheaf~$E$ of~$\O_Y$-modules can
be regarded as a sheaf over~$I(Y)$. Under this equivalence, the morphism~$f : Y
\to X$ corresponds to the unique morphism~$I(Y) \to \mathrm{pt}$ to the
internal one-point locale. Hence it makes sense to construct, from the internal
point of view of~$\Sh(X)$, the sheaf cohomology~$H^n(I(Y), E)$ of~$E$.

Usually one would not expect an internal construction which depends on
arbitary choices to yield a globally-defined sheaf on~$X$ -- following the
definition of the stack semantics we only obtain a family of sheaves defined on
members of some open covering of~$X$; but we verify in
Theorem~\ref{thm:higher-direct-images-as-internal-sheaf-cohomology} that in our
case, it does, and that the resulting sheaf coincides with the~$n$-th higher
direct image of~$E$.

\begin{lemma}\label{lemma:notions-of-internal-injectivity}
Let~$Y$ be a ringed locale over a locale~$X$. Let~$I$ be a sheaf of modules on~$Y$.
Assuming Zorn's lemma in the metatheory, the following statements are equivalent:
\begin{enumerate}
\item $I$ is an injective sheaf of modules.
\item From the point of view of~$\Sh(Y)$, $I$ is an injective module.
\item From the point of view of~$\Sh(X)$, $I$ is an injective module from the
point of view of~$\Sh(I(Y))$.
\item From the point of view of~$\Sh(X)$, $I$ is an injective sheaf of modules
on~$I(Y)$.
\end{enumerate}
\end{lemma}

\begin{proof}The equivalence of the first two statements is by
Theorem~\ref{thm:injectivity-external-internal}. The equivalence~$\text{(2)}
\Leftrightarrow \text{(3)}$ is by the idempotency of the stack semantics:
$\Sh(Y) \models \varphi$ if and only if~$\Sh(X) \models (\Sh(I(Y)) \models
\varphi)$. (Shulman stated and proved a restricted version of this idempotency property
in his original paper on the stack semantics~\cite[Lemma~7.20]{shulman:stack-semantics}.
A proof of the general case is slightly less
accessible~\cite[Lemma~1.20]{blechschmidt:master}.) The equivalence~$\text{(3)}
\Leftrightarrow \text{(4)}$ is by interpreting
Theorem~\ref{thm:injectivity-external-internal} internally to~$\Sh(X)$. This
requires Zorn's lemma to hold internally to~$\Sh(X)$; this is indeed the case
since we assume Zorn's lemma in the metatheory and since the validity of Zorn's
lemma passes from the metatheory to localic
toposes~\cite[Proposition~D4.5.14]{johnstone:elephant}.
\end{proof}

\begin{thm}\label{thm:higher-direct-images-as-internal-sheaf-cohomology}
Let~$f : Y \to X$ be a ringed locale over a locale~$X$. Let~$E$ be a sheaf of modules over~$Y$.
Assuming the axiom of choice in the metatheory, the expression~``$H^n(I(Y), E)$''
of the internal language of~$\Sh(X)$ denotes a globally-defined sheaf
on~$X$, and this sheaf coincides with~$R^n f_*(E)$.\end{thm}

\begin{proof}By Lemma~\ref{lemma:notions-of-internal-injectivity} and by the
fact that every sheaf of modules over~$Y$ admits an injective resolution, every sheaf
of modules over~$I(Y)$ admits an injective resolution from the point of view
of~$\Sh(X)$. Hence we can, internally to~$\Sh(X)$, carry out the construction of~$H^n(I(Y), E)$.
Externally, this yields an open covering of~$X$ such that we have, for each
member~$U$ of that covering
\begin{itemize}
\item a sheaf~$M$ on~$U$,
\item a module structure on~$M$,
\item a resolution~$0 \to E|_{f^{-1}U} \to I^\bullet$ by sheaves of modules which are
internally and hence externally injective and
\item data exhibiting~$M$ as the~$n$-th cohomology
of~$(f|_{f^{-1}U})_*(I^\bullet)$.
\end{itemize}
On intersections of such opens~$U$ and~$U'$, there is exactly one
isomorphism~$M|_{U \cap U'} \to M'|_{U \cap U'}$ of sheaves of modules induced by
a morphism of resolutions which is compatible with the identity on~$E$.
Hence the cocycle condition for these isomorphisms is satisfied, ensuring that
the individual sheaves~$M$ glue to a globally-defined sheaf of modules on~$X$.
(The individual injective resolutions need not glue to a global injective
resolution.)

The claim that this sheaf coincides with~$R^n f_*(E)$ follows from the fact
that we can pick as internal resolution of~$E$ (considered as a sheaf
over~$I(Y)$), the particular injective resolution of~$E$ (considered as a sheaf
over~$Y$) used to define~$R^n f_*(E)$.
\end{proof}
